\documentclass[letterpaper,10pt,notopicbreak,titleabove,plain]{simplecv}
\usepackage{ifpdf}
%\usepackage{graphicx}
%\usepackage{color}
%\usepackage{pgf}
\ifpdf
\usepackage[pdftex,bookmarks=false]{hyperref}
\else
\usepackage[hypertex]{hyperref}
\fi
\setlength{\oddsidemargin}{0pt}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-37pt}
\setlength{\textheight}{9in}
\renewcommand{\topicmargin}{0.07\columnwidth}

\leftheader{%
%2342 California St, Apt 11\\
%Mountain View CA 94040
James H Clark Center, Rm S245\\
318 Campus Drive W\\
Stanford CA 94305
}

\rightheader{%
TEL: +1-618-201-6128\\
%TEL: +44-(0)2380-732904\\
\href{mailto:sulahiri@fstanford.edu.delete.this}{\texttt{sulahiri@fstanford.edu}}
%\href{mailto:subhylahiri@gmail.com}{\texttt{subhylahiri@gmail.com}}
}

\title{Subhaneil Lahiri}


\begin{document}
\pagestyle{empty}
\maketitle


\section{Postdoctoral positions}
\begin{topic}
\item[Ganguli Lab:]  (Postdoctoral Fellow, Stanford University, Stanford CA, March 2012 -- present). Computational and theoretical Neurobiology.

\item[Samuel Lab:]  (Biophysics, Harvard University, Cambridge MA, July 2009 -- August 2011). Performing quantitative analysis of brain and behavior in the \emph{Drosophila} larva.
\end{topic}

\section{Education}
\begin{topic}
  \item[Harvard University, Cambridge MA:] Ph.D., Physics, June 2009. \\
  Ph.D.\ Thesis: Black holes from fluid mechanics. Advisor: Shiraz Minwalla.% GPA: 3.9/4.0.

  \item[Tata Institute, Mumbai, India:] Visiting student, Spring 2006 -- Spring 2007. Research in theoretical physics.
 %     \\
 %     Additional reading group: Stochastic calculus and mathematical finance.

  \item[Oxford University, Oxford, UK:] M.Phys., 1st class honors, Merton college, 2003. \\
  M.Phys.\ Project: Monte Carlo simulations of galactic heating. %Overall score (GPA): 92\%.

  %\item[Colchester Royal Grammar School, Colchester UK:] A-level (1999): Physics (A), Chemistry (A), Mathematics (A), Further Mathematics (A) and General Studies (A). (S-level: Physics (Distinction), Chemistry (Distinction)). GCSE (1997): Mathematics (A*), Physics (A*), Chemistry (A*), Biology (A*), Geography (A*), German (A*), French (A), Technology (A), English Language (A), English Literature (B), Art and Design (B).
\end{topic}

\section{Published articles}
\nocite{Lahiri2013synapse,Advani2013repRev,Lahiri2011,Bhattacharya:2009gm,Bhattacharyya:2007vs,Lahiri:2007ae,Biswas:2006tj}
\bibliographystyle{utcaps_sl}
\bibliography{adscft,larva,neuro}

%\section{Other research experience}
%\begin{topic}
%  \item[Ph.D.\ Thesis:] Black holes from fluid mechanics (Harvard, 2009).
%
%  \item[M.Phys.\ Project:] Monte Carlo simulations of galactic heating (Oxford, 2002).
%
%  \item[Vacation Project:] Simulations of liquid crystal defect dynamics (Oxford, 2001).
%\end{topic}

\section{Awards and Fellowships}
\begin{topic}
   \item   Winner of the Neural Information Processing Systems Outstanding Paper Award, given to 3 papers out of 1420 submissions (Neural Information Processing Systems, December 2013).
   \item   Stanford Bio-X Genentech Fellowship (Stanford University, Stanford CA, 2013).
   \item   Certificate of distinction in teaching (Harvard University, Cambridge MA, Spring 2008).
   \item   Vineer Bhansali Graduate Travel Fellowship (Harvard University, Cambridge MA, 2006).
   \item   James Mills Pierce Fellowship (Harvard University, Cambridge MA, 2003).
   \item   Scott Prize for best performance in M.Phys.\ examination (Oxford University, Oxford, UK, 2003).
   \item   Commendation for prelims practical work (Oxford University, Oxford, UK, 2000).
\end{topic}

\section{Teaching experience}
\begin{topic}
  \item[SERC School Tutor:] Taught at an intensive school in high energy theoretical physics for graduate students in India (Hyderabad University, Hyderabad, India, 2007).

  \item[Harvard University Teaching Fellow:] Taught physics classes for undergraduates in physics and other departments (Harvard University, Cambridge MA, Fall 2004 -- Fall 2005, Spring 2008 -- Fall 2010).

  \item[Harvard Summer Teaching Fellow:] Taught physics daily to high school students and undergraduates from various colleges (Harvard University, Cambridge MA, 2005).
\end{topic}




%\section{Employment}
%\begin{topic}
%	\item Part time waiter at Wivenhoe House Hotel, 7/00 -- 9/00, 12/00 -- 1/01 and 8/01 -- 9/01.
%
%	\item Sales Assistant at Matalan Discount Store from 7/00 -- 8/00.
%
%    %\item Part time Physics Laboratory Assistant at Colchester Royal Grammar School, 5/98 -- 5/99.
%\end{topic}
%
%
%\section{Leadership}
%\begin{topic}
%	\item Organizer of the Graduate String Theory Seminars, Harvard, 2005 -- 06.
%	\item Publicity Officer of the Oxford Physics Society, 3/00 -- 6/01.
%%	\item Stage Manager for school production of Oliver! in 3/99.
%\end{topic}

\section{Conference presentations}
\begin{topic}
  \item[Neural Information Processing Systems (NIPS):] Lake Tahoe NV. Dec 5-8, 2013 .\\
  Poster: A memory frontier for complex synapses.
  \item[Sloan-Swartz Centers for Theoretical Neurobiology Annual Meeting:] Brandeis University, Waltham MA. July 26-28, 2013 .\\
  Talk: Understanding impaired learning with enhanced plasticity.
  \item[Computational and Systems Neuroscience (Cosyne):] Salt Lake City UT. February 28 - March 3, 2013.\\
  Poster: A general theory of learning and memory with complex synapses.
  \item[Sloan-Swartz Centers for Theoretical Neurobiology Annual Meeting:] UC San Diego, La Jolla CA. June 27–29, 2012.\\
  Poster: Learning and memory with complex synapses.
%  \item[Sloan-Swartz Centers for Theoretical Neurobiology Annual Meeting:] Harvard University, Cambridge MA. July 25--28, 2009.
%  \item[Monsoon Workshop on String Theory:] TIFR, Mumbai, India. July 7--August 3, 2008.
%  \item[TASI 2007:] University of Colorado, Boulder CO. May 28–-June 22, 2007.
  \item[1st Asian Winter School on String Theory:] KIAS, Seoul, S.~Korea. January 8--19, 2007.\\
    Talk: Giant gravitons and the supersymmetric states of N= 4 Yang-Mills.
%  \item[Indian Strings Meeting:] Puri, India. December 12--19, 2006.
%  \item[Strings 2006:] Beijing, China. June 19--24 2006.
%  \item[Strings 2005:] Toronto, Canada. July 10--16, 2005.
%  \item[Strings, Gravity \& Cosmology Summer School:] UBC, Vancouver, Canada. August 2004.
%  \item[Introductory School on Recent Developments in Supersymmetric Gauge Theories:] ICTP, Trieste, Italy. June 14--25, 2004.
\end{topic}


%\section{Personal information}
%\begin{topic}
%  \item[Citizenship:] British.
%
%  \item[Date of birth:] February 4, 1981.
%\end{topic}


\end{document}



