\documentclass[letterpaper,10pt,notopicbreak,titleabove,plain]{simplecv}
\usepackage{ifpdf}
\setlength{\oddsidemargin}{0pt}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-37pt}
\setlength{\textheight}{9in}
\renewcommand{\topicmargin}{0.10\columnwidth}
%\usepackage{graphicx}
%\usepackage{color}
%\usepackage{pgf}
%\usepackage{hyperref}
\ifpdf
\usepackage[pdftex,bookmarks=false]{hyperref}
\else
\usepackage[hypertex]{hyperref}
\fi

%\leftheader{Jefferson Physical Laboratory\\
%17 Oxford Street\\
%Cambridge MA 02138, USA}
%
%\rightheader{TEL: +1-618-201-6128\\
%\href{mailto:slahiri@fas.harvard.edu}{\texttt{slahiri@fas.harvard.edu}}}

\leftheader{TEL: 618-201-6128}
\rightheader{\href{mailto:slahiri@fas.harvard.edu.delete.this}{\texttt{slahiri at fas.harvard.edu}}}

\title{Subhaneil Lahiri%
\\ {\normalfont\normalsize 15 Saint Paul St.\ Apt.\ 2, Cambridge MA 02139}}

\begin{document}
\pagestyle{empty}
\maketitle

\section{Current occupation}
\begin{topic}
\item[Postdoctoral Research Fellow:] Samuel Lab (Biophysics, Harvard University, Cambridge MA, July 2009 -- present). Performing quantitative analysis of brain and behavior in the \emph{Drosophila} larva.
\end{topic}

\section{Education}
\begin{topic}
  \item[Harvard University, Cambridge MA:] Ph.D., Physics, June 2009. \\
  Ph.D.\ Thesis: Black holes from fluid mechanics. Advisor: Shiraz Minwalla.% GPA: 3.9/4.0.

  \item[Tata Institute, Mumbai, India:] Visiting student, Spring 2006 -- Spring 2007. Research in theoretical physics.
 %     \\
 %     Additional reading group: Stochastic calculus and mathematical finance.

  \item[Oxford University, Oxford, UK:] M.Phys., 1st class honors, Merton college, 2003. \\
  M.Phys.\ Project: Monte Carlo simulations of galactic heating. %Overall score (GPA): 92\%.
\end{topic}

%\section{Research}
%\begin{topic}
%  \item[Ph.D.\ Thesis:] Black holes from fluid mechanics (Harvard, in progress).
%
%  \item[M.Phys.\ Project:] Monte Carlo simulations of galactic heating (Oxford, 2002).
%
%  \item[Vacation Project:] Simulations of liquid crystal defect dynamics (Oxford, 2001).
%\end{topic}

\section{Teaching experience}
\begin{topic}
  \item[SERC School Tutor:] Taught at an intensive school in high energy theoretical physics for graduate students in India (Hyderabad University, Hyderabad, India, 2007).

  \item[Harvard University Teaching Fellow:] Taught physics classes for undergraduates in physics and other departments (Harvard University, Cambridge MA, Fall 2004 -- Fall 2005, Spring 2008 -- Fall 2010).

  %\item[Harvard Summer Teaching Fellow:] Taught physics daily to high school students and undergraduates from various colleges (Harvard University, Cambridge MA, 2005).
\end{topic}

\section{Additional experience}
\begin{topic}
	\item Part time waiter at Wivenhoe House Hotel, Colchester, UK, 7/00 -- 9/00, 12/00 -- 1/01, 8/01 -- 9/01 and 7/03 -- 8/03.

	\item Sales Assistant at Matalan Discount Store, Colchester, UK, 7/00 -- 8/00.

%    \item Part time Physics Laboratory Assistant at Colchester Royal Grammar School, 5/98 -- 5/99.
\end{topic}

\section{Awards and Fellowships}
\begin{topic}
   \item   Certificate of distinction in teaching (Harvard University, Cambridge MA, Spring 2008).
   \item   Vineer Bhansali Graduate Travel Fellowship (Harvard University, Cambridge MA, 2006).
   \item   James Mills Pierce Fellowship (Harvard University, Cambridge MA, 2003).
   \item   Scott Prize for best performance in M.Phys.\ examination (Oxford University, Oxford, UK, 2003).
%   \item   Commendation for prelims practical work (Oxford, 2000).
\end{topic}

\section{Leadership experience}
\begin{topic}
	\item Organizer of the Graduate String Theory Seminars at Harvard University, arranging speakers for a weekly seminar series and managing a budget of \$4500 per year (Harvard University, Cambridge MA, 2005 -- 06).
	\item Co-founder and Publicity Officer of the Oxford University Physics Society, marketing events via posters, announcements and a website (Oxford University, Oxford, UK, 3/00 -- 6/01).
%	\item Stage Manager for school production of Oliver! in 3/99.
\end{topic}



%\section{Citizenship}
%British.

%\section{Publications}
%\nocite{Bhattacharya:2009gm,Bhattacharyya:2007vs,Lahiri:2007ae,Biswas:2006tj}
%\bibliographystyle{utcaps}
%\bibliography{adscft}


\end{document}



